---
title: "EDA Project"
author: "Blake Bell, Cassandra Gibbs, Iain Kirsch, Tyler O'Toole, Rushma Khatri Subedi"
date: "`r Sys.Date()`"
output: 
  ioslides_presentation:
    widescreen: TRUE
    fig_height: 5
    fig_width: 10
    df_print: paged
runtime: shiny
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = 'png')
library(tesseract)
library(magick)
library(tidycensus)
library(tidyverse)
library(tigris)
library(sf)
library(viridis)
library(shiny)
library(shinydashboard)
library(rsconnect)
library(png)
library(knitr)
library(leaflet)
library(rgdal)
library(stringr)
library(RColorBrewer)
library(colorspace)
library(scales)
library(corrplot)
library(kableExtra)
library(jtools)
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)

source("data.Rdmpd")
source("data2.Rdmpd")
source("data3.Rdmpd")
source("data4.Rdmpd")
data <- read.csv("projectdata.csv")
data$Rape_Rate = data$Rapes/data$Total_Pop
data$Rob_Rate = data$Robberies/data$Total_Pop
data$Murder_Rate = data$Murders/data$Total_Pop
data2<- data %>% arrange((Woman_LFPR))
data$MSA<- as.character(data$MSA)
rownames(data)<- NULL
g  <- lm(Woman_LFPR ~ Car_Commute + Index + Transit_Commute + Rape_Rate + Rob_Rate + Murder_Rate +
           Population_Usage*Income_Difference , data = data)
# Correlation Matrix
correlation_data <- data[, c(3:18)]
res <- cor(correlation_data)
round(res,2)
head(res,5)
```

``` {r echo=FALSE}
metro_areas<- unique(female_lforce$metro_name)
datalist<- list()
lf<- female_lforce$estimate/female_pop$estimate
datamap<- cbind(female_lforce, lf)
datamap2<- st_as_sf(datamap, sf_column_name= datamap$geometry)
pal <- colorQuantile(palette = "viridis", domain = datamap2$lf, n=10)
df<- data %>% 
  select(-X)

lfprwomangrapher<- function(x) {
  localwomenlforce<- female_lforce %>% 
    filter(metro_name== x)
  lfwomenlocal<- sum(localwomenlforce$estimate)
  
  localtotallforce<- total_lforce %>% 
    filter(metro_name==x)
  lftotallocal<- sum(localtotallforce$estimate)
  
  localpopwomen<- female_pop %>%
    filter(metro_name==x)
  popwomenlocal<- sum(localpopwomen$estimate)
  
  localpoptotal<- total_pop %>%
    filter(metro_name==x)
  poptotallocal<- sum(localpoptotal$estimate)
  
  #Labor Force Participation Rate for San Francisco
  lftotallocal/poptotallocal
  lfwomenlocal/popwomenlocal
  
  LFPR<- localwomenlforce$estimate/localpopwomen$estimate
  women_population_local<- localpopwomen$estimate
  LFPR_Total_local<- localtotallforce$estimate/localpoptotal$estimate
  total_population_local<- localpoptotal$estimate
  local_total<- cbind(localtotallforce, LFPR_Total_local, total_population_local, LFPR, women_population_local)
  
  # Women
  ggplot(local_total, aes(fill = LFPR, color = LFPR)) + 
    geom_sf() + 
    coord_sf(crs = 26910) + 
    theme_minimal() + 
    theme(aspect.ratio = 1) + 
    scale_fill_viridis() + 
    scale_color_viridis()
}
lfprtotalgrapher<- function(x) {
  localwomenlforce<- female_lforce %>% 
    filter(metro_name== x)
  lfwomenlocal<- sum(localwomenlforce$estimate)
  
  localtotallforce<- total_lforce %>% 
    filter(metro_name==x)
  lftotallocal<- sum(localtotallforce$estimate)
  
  localpopwomen<- female_pop %>%
    filter(metro_name==x)
  popwomenlocal<- sum(localpopwomen$estimate)
  
  localpoptotal<- total_pop %>%
    filter(metro_name==x)
  poptotallocal<- sum(localpoptotal$estimate)
  
  #Labor Force Participation Rate for San Francisco
  lftotallocal/poptotallocal
  lfwomenlocal/popwomenlocal
  
  LFPR_Women_local<- localwomenlforce$estimate/localpopwomen$estimate
  women_population_local<- localpopwomen$estimate
  LFPR<- localtotallforce$estimate/localpoptotal$estimate
  total_population_local<- localpoptotal$estimate
  local_total<- cbind(localtotallforce, LFPR, total_population_local, LFPR_Women_local, women_population_local)
  
  # Graphing MSA Labor Force Participation Rate
  par(mfrow=c(2,1))
  ggplot(local_total, aes(fill = LFPR, color = LFPR)) + 
    geom_sf() + 
    coord_sf(crs = 26910) + 
    theme_minimal() + 
    theme(aspect.ratio = 1) + 
    scale_fill_viridis() + 
    scale_color_viridis()
}
```

## Question
<div class="blue">
- Does crime or the quality of public transportation available affect women's labor force participation rate?
- We studied Metropolitan Statistical Areas in the United States and merged datasets together to get data for 16 MSAs
- Our dataset is able to answer this question for the largest American cities, but should not be extrapolated to smaller or foreign cities
</div>

# Data Import & Transformation

## Datasets
<div class="blue">
- ACS Census Data on Labor Force Participation and Population on the Census Tract level off an API
- Kaggle Crime Data for the United States on the County level
- SmartAsset's Public Transportation Image on the City level
  + https://dr5dymrsxhdzh.cloudfront.net/blog/images/a18525/2016/02/public_transportation_1_table.png

</div>

## Data
<div class="blue">
- Wrangling involved releveling, and spatial joins for ACS data
- Relevelling and bringing up crime data to the MSA level
- ACS Dataset had 158,076 observations of 7 variables
- Crime Data had 3,136 observations of 24 variables
- Transportation Data had 25 observations of 7 variables
</div>

## Final Dataset

``` {r echo=FALSE}
df %>% 
  kable() %>%
  kable_styling(c( "bordered", "condensed", "reponsive")) %>%
  add_header_above(c(" ", "Census Data" = 2, "Crime Data" = 4, "Transportation Data" = 10)) %>% 
  row_spec(, row=1:16, bold=TRUE, background="white") %>% 
  row_spec(, row=0, bold=TRUE, background="white") %>%
  scroll_box(width = "800px", height = "500px")
```

## Methodology
<div class="blue">
- Basic OLS Regression on 16 Major American Cities
- Limitations on Transportation data led to small MSA level sample size
- Small number of observations reduces number of variables that can be used effectively

</div>

# Packages

## Packages Used
<div class="blue">
- tidycensus: Used to pull ACS data
- tesseract: Used to run OCR on images
- magick: Image processing in R
- tidyverse (dplyr, ggplot2): Data manipulation and visualization
- tigris, sf: Used for Census TIGER shapefiles, Simple features for spatial data
- knitr, kableExtra: tables and slidebox
- viridis, RColorBrewer, scales, colorspace: Color schemes for graphs
- shiny, shinydashboard, rsconnect: Used for interactive graphs
- leaflet, rgdal, stringr: Used for large maps of the US
- corrplot, jtools: Correlation plot to look at variables, effect plots
</div>

# Exploratory Data Analysis

## MSAs Studied

``` {r echo=FALSE}
leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addMarkers(lng = -122.41, lat = 45.31, popup = "Portland") %>%
  addMarkers(lng = -122.26, lat = 37.47, popup = "San Francisco") %>%
  addMarkers(lng = -89.40, lat = 43.07, popup = "Madison") %>%
  addMarkers(lng = -73.58, lat = 40.47, popup = "New York") %>%
  addMarkers(lng = -76.38, lat = 39.18, popup = "Baltimore") %>%
  addMarkers(lng = -111.54, lat = 40.46, popup = "Salt Lake City") %>% 
  addMarkers(lng = -81.37, lat = 41.28, popup = "Cleveland") %>%
  addMarkers(lng = -71.50, lat = 42.21, popup = "Boston") %>%
  addMarkers(lng = -122.20, lat = 47.37, popup = "Seattle") %>%
  addMarkers(lng = -87.37, lat = 41.50, popup = "Chicago") %>%
  addMarkers(lng = -93.14, lat = 44.59, popup = "Minneapolis") %>%
  addMarkers(lng = -75.10, lat = 39.57, popup = "Philadelphia") %>%
  addMarkers(lng = -77.02, lat = 38.53, popup = "Washington DC") %>%
  addMarkers(lng = -77.29, lat = 37.33, popup = "Richmond") %>%
  addMarkers(lng = -78.50, lat = 42.55, popup = "Buffalo") %>%
  addMarkers(lng = -105.00, lat = 39.45, popup = "Denver")
```

## Best Areas for Women's Labor Force Participation

``` {r echo=FALSE}
datamap2 %>%
  st_transform(crs = "+init=epsg:4326") %>%
  leaflet(width = "100%") %>%
  addProviderTiles(provider = "Esri.WorldImagery") %>%
  addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
              stroke = FALSE,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal(lf)) %>%
  addLegend("bottomright", 
            pal = pal, 
            values = ~ lf,
            title = "LFPR Deciles",
            opacity = 1)

```

## Labor Force Participation

```{r echo=FALSE}
ui <- dashboardPage(skin = "blue",
                    dashboardHeader(title= "LFPR by MSA", titleWidth = 200),
                    dashboardSidebar(disable=TRUE),
                    dashboardBody(
                      fillPage(padding = 0,
           box(title="Labor Force Participation Rate", id="normal", solidHeader = TRUE, status = "info",
                    plotOutput("plot1", height = 250)), 
            
            box(title="MSA to Examine", background = "black",
                    selectInput("variable", "MSA:", choices= sort(metro_areas), selected="San Francisco-Oakland-Hayward, CA"))
            )))


server <- function(input, output) {
  
   
    output$plot1<- renderPlot({
    par(mfrow=c(1,1))
    par(mar = c(4, 5, 4, 4)) 
    lfprtotalgrapher(input$variable)
    #add_legend("bottomright", legend=c("State Recession", "National Recession"),
               #col=c("grey", "red"), pch=15, bty='n', horiz=TRUE)
    })

   }


shinyApp(ui=ui, server=server)

```

## Women's Labor Force Participation

```{r echo=FALSE}
ui <- dashboardPage(skin = "blue",
                    dashboardHeader(title= "Women's LFPR by MSA", titleWidth = 200),
                    dashboardSidebar(disable=TRUE),
                    dashboardBody(
                      fillPage(padding = 0,
           box(title="Labor Force Participation Rate", id="normal", solidHeader = TRUE, status = "info",
                    plotOutput("plot1", height = 250)), 
            
            box(title="MSA to Examine", background = "black",
                    selectInput("variable", "MSA:", choices= sort(metro_areas), selected="San Francisco-Oakland-Hayward, CA"))
            )))


server <- function(input, output) {
  
   
    output$plot1<- renderPlot({
    par(mfrow=c(1,1))
    par(mar = c(4, 5, 4, 4)) 
    lfprwomangrapher(input$variable)
    #add_legend("bottomright", legend=c("State Recession", "National Recession"),
               #col=c("grey", "red"), pch=15, bty='n', horiz=TRUE)
    })

   }


shinyApp(ui=ui, server=server)

```

## Study of LFPR vs. Women's LFPR

```{r echo=FALSE}
ggplot(data, aes(x=Total_LFPR, y=Woman_LFPR, col = MSA))+
  geom_point(size=3, alpha = .75)+
  ggtitle("Labor Force Participation Rates")+
  labs(x= "Total LFPR", y = "Women's LFPR") +
  scale_y_continuous(labels = percent) +
  scale_x_continuous(labels = percent) +
  theme_dark()

```

## Crime Rate vs. Women's LFPR

```{r echo=FALSE}
ggplot(data, aes(x=Crime_Rate_Per_100k, y=Woman_LFPR, col = MSA))+
  geom_point(size=3, alpha = .75)+
  ggtitle("Crime Rates and Women's LFPR")+
  labs(x= "Crime Per 100k", y = "LFPR") +
  scale_y_continuous(labels = percent) +
  theme_dark()
```

## Women's LFPR by MSA

```{r echo=FALSE}
ggplot(data, aes(x=reorder(MSA, Woman_LFPR), y=Woman_LFPR, fill=MSA))+
  geom_bar(stat = "identity")+
  ggtitle("Women's LFPR by MSA")+
  labs(x="", y = "")+
  scale_y_continuous(labels = percent) +
  scale_fill_viridis(discrete = TRUE) +
  theme_dark()+
  theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip() +
  guides(fill=FALSE)
```

## Transportation Quality Index

```{r echo=FALSE}
ggplot(data, aes(x=reorder(MSA, Index), y=Index, fill=MSA))+
  geom_bar(stat = "identity")+
  ggtitle("Index by MSA")+
  labs(x="", y = "")+
  scale_y_continuous() +
  scale_fill_viridis(discrete = TRUE) +
  theme_dark()+
  theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip() +
  guides(fill=FALSE)
```

## Correlation Plot
```{r echo=FALSE}
corrplot(res, method = "circle")
```

# Results

## Regression Results

<div class="blue">
- When the Transportation Index increases by 1, women's participation increases by 0.007 (units are in percentage points)
- When the robbery rate goes up by 1 percentage point, women's participation decreases by 0.4 percentage points
- When the average time it takes to commute by car increases by 1 minute, the labor force participation rate for women decreases by 0.014 percentage points
- The model explains 96.11% of the variation in the women's labor force participation rate
</div>

## Mixed Results

```{r echo=FALSE}
ui <- dashboardPage(skin = "blue",
                    dashboardHeader(title= "Effectiveness of Analysis", titleWidth = 300),
                    dashboardSidebar(disable=TRUE),
                    dashboardBody(
                      fillPage(padding = 0,
           box(title="Effect Plot for Robbery Rate", id="normal", solidHeader = TRUE, status = "info",
                    plotOutput("plot1", height = 250)),
           box(title="Effect Plot for Transportation Index", id="normal", solidHeader = TRUE, status = "info",
                    plotOutput("plot2", height = 250))
            )))


server <- function(input, output) {
  
   
    output$plot1<- renderPlot({
    effect_plot(g, pred = Rob_Rate, interval = TRUE, plot.points = TRUE, x.label = "Robbery Rate", 
            y.label = "Women's Labor Force Participation Rate")
    })
    output$plot2<- renderPlot({
    effect_plot(g, pred = Index, interval = TRUE, plot.points = TRUE, x.label = "Robbery Rate", 
            y.label = "Women's Labor Force Participation Rate")
    })

   }


shinyApp(ui=ui, server=server)

```

## Summary of Results

<div class="blue">
- The Murder and Robbery Rates, in addition to the Index (Overall Transportation Score) are all significant at the 5% level
- The results are mixed, and in order to better understand the question, data at the county or census tract level for transportation would be required
- Individual Crime Rates are more important to labor force participation for women than the overall crime rate
- Better transportation does appear to improve labor force participation, but likely not as much as our model suggests
</div>

## Works Cited

<div class="blue">
- "Data Access and Dissemination Systems (DADS)." American FactFinder API, United States Census Bureau, 5 Oct. 2010, factfinder.census.gov/faces/nav/jsf/pages/searchresults.xhtml?refresh=t.
- "US Crime Data." Kaggle, www.kaggle.com/datasets.
- Wallace, Nick. "The Best Cities for Public Transportation." SmartAsset, SmartAsset, 20 Sept. 2018, smartasset.com/mortgage/best-cities-for-public-transportation.
</div>

## Questions?

```{r echo=FALSE}
ggplot(data, aes(x=reorder(MSA, Index), y=Index, fill=MSA))+
  geom_bar(stat = "identity")+
  ggtitle("Index by MSA")+
  labs(x="", y = "")+
  scale_y_continuous() +
  scale_fill_viridis(discrete = TRUE) +
  theme_dark()+
  theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip() +
  guides(fill=FALSE)
```





